#Chapter template

##Overview

(Write briefly about this region, listing the states included within it, the range of seats/districts for each state, and the general partisan landscape (lean Republican vs. Democratic). The echo and message parameters below keeps the code itself out of the document.)

```{r, echo=FALSE, message=FALSE}
#Here's some example code
library(tidyverse)
library(sf)

proj_region="Midwest Region" #Adjust to fit your group

district_data<-read_csv("https://github.com/jshannon75/geog4300/raw/master/Labs/Final_assignment/districts_data_all.csv") %>%
  mutate(year=as.character(year))

library(ggridges)
districts_select<-district_data %>%
  filter(Region==proj_region) %>%
  group_by(year) %>%
  mutate(median_dem=mean(pct_dem),
         pct_dem_adj=ifelse(uncontest_r==1,66,ifelse(uncontest_d==1,36,pct_dem))) %>%
  filter(total_vote!=-99)

ggplot(districts_select,aes(x=pct_dem_adj,y=st_abbr))+
  geom_density_ridges(scale=3)+
  theme_ridges()+
  geom_vline(aes(xintercept = median_dem),col="red")+
  facet_wrap(~year)+
  xlab("Pct. voting Democractic")+
  ylab("")

```

##Descriptive analysis

(Include both ridge plots and maps here showing your independent variable. Interpret what both tell you. You may also want to include state median values/IQR to help describe the patterns you see. Note that here the spatial data files needs to be loaded in a chunk with "include=FALSE" to surpress the loading message.)

```{r, include=FALSE}
##Some sample maps
library(tmap)
districts<-st_read("https://github.com/jshannon75/geog4300/raw/master/Labs/Final_assignment/districts_all.geojson",stringsAsFactors=FALSE) 
districts_join<-left_join(districts,districts_select) %>%
  filter(st_abbr!="AK" & st_abbr!="HI")

states<-st_read("https://github.com/jshannon75/geog4300/raw/master/Labs/Final_assignment/USstates_48.geojson") %>% filter(Region==proj_region)
```

```{r, echo=FALSE, message=FALSE}
districts_1990<-districts_join %>% filter(year=="1990" & Region==proj_region)
districts_2015<-districts_join %>% filter(year=="2015" & Region==proj_region)

map1<-tm_shape(districts_1990,projection="2163") +
  tm_polygons("pct_dem_adj", breaks=c(0,40,45,50,55,70,100),border.alpha=0)+
  tm_shape(states) + tm_borders()

map2<-tm_shape(districts_2015,projection="2163") +
  tm_polygons("pct_dem_adj", breaks=c(0,40,45,50,55,70,100),border.alpha=0)+
  tm_shape(states) + tm_borders()

tmap_arrange(map1,map2)
```

##Models

(Include models for both years and scatterplots/fit lines. Interpret the coefficients and measures of model strength, focusing on direction, magnitude, and significance.)

##Diagnostics

(Look at model residuals using density plots and maps as shown in the assignment. Does your model show any warning signs based on this output).

##Summary
(What did this analysis tell you about the research questions: the relationship between your variable and voting and changes in the variable and its relationship to voting across time periods?)